Generative Adversarial Networks

### 2017-12-03 cs6735/project/draft.tex
small fix
lots of editing

\begin{abstract}
    Deep generative networks have often fallen short, especial when compared to discriminative models. Since their conception in 2014, generative adversarial networks have seen exponential growth in research. These networks offer promising results for both unsupervised learning and generative models. This research paper presents an overview of generative adversarial networks. Specifically, we will examine the motivation behind GANs, their theoretical justification, the current state of GANs and the research frontiers that have yet to be explored.
\end{abstract}

\section{Introduction}
Deep learning have made astounding progress within the past decade. Discriminative models have recently surpassed the abilities of human within the domain of pattern recognition \citep{2014arXiv1404.7828S}. These successes can attributed to vast, high dimension datasets in conjunction with large neural networks using linear activation functions, dropout regularization techniques and backpropagation to update the parameters \citep{2014arXiv1406.2661G}. However, deep learning possesses many more ambitious goals. Until recently, success has mostly been seen with supervised classifiers; however, deep generative models now competitively rival their discriminative counterparts.

Generative models learn the joint probability distribution $p(x,y)$ of an input $x$ and label $y$ whereas discriminative models directly learn the conditional probability $p(y|x)$. Generative models may be used as classifiers using Bayes rules to calculate the conditional probability $p(x,y)$ which can then be used to make predictions \citep{NIPS2001_2020}. Depending on the model, knowledge of the probability distribution can be created both explicitly and implicitly \citep{Goodfellow-et-al-2016}. Those which do not directly model a probability distribution offer mechanisms which require implicit knowledge of the underlying distribution, such as creating a sample from that distribution \citep{Goodfellow-et-al-2016}. As humans, we are able to understand the world around us with tremendous precision. It is easy to underestimate the complexity of the data we process to accomplish this feat. Although progress within the machine learning field is rapidly advancing, computers still have limited understanding of the data the process. Generative models, especially deep generative models, offer promising results towards this goal \citep{genmodelingopenai}.

\begin{figure}
\centering
\includegraphics[scale=0.7]{pdf}
\caption{The process density estimation of one-dimensional data and a Gaussian distribution \citep{2017arXiv170100160G}. Generative models take a dataset $D$, sourced from a distribution $p_{data}$, a create an estimate of that distribution $p_{model}$.}
\label{fig:pdf}
\end{figure}

The process of training a generative model is very similar to a discriminative model. Using large amounts of data collected from a specific domain, we train the model to generate data from that domain. As these models have fewer parameters than that the number of data samples, they are forced to internalize some representation of the data. Although, unlike supervised training, there exists no explicit desired output. The problem becomes defining a cost function which forces the generative model to produce data more like that of the domain.

Generative adversarial networks (GANs), first introduced in 2014, offer a new framework for estimating generative models with use of an adversarial process \citep{2014arXiv1406.2661G}. These models offer a clever approach to solving the aforementioned problem of explicitly defining a cost function. Rather than training a single model, a discriminator is introduced. These two networks are pitted against each other in a minimax game. The generative model $G$ attempts to produce data that resembles that of the training set while the discriminative model $D$ attempts to classify whether or not a sample was real of generated. To train this network, $G$ attempts to maximize the error rate of $D$ whereas $D$ attempts to minimize it. The goal of the GAN algorithm is to reduce the distance between the data and generated probability distributions. Backpropagation is used to update the parameters and to train each model. In the ideal case, $D$ is unable to distinguish the generated samples from the real samples and produces an error rate of $\frac{1}{2}$ \citep{2014arXiv1406.2661G}.

\section{Motivations}
There exists several compelling reasons for studying generative modeling \citep{2017arXiv170100160G}. These models are extremely useful when your goal is understand the underlying distribution parameters. For example, joint probability distributions of high dimensional data is relevant to both applied math and engineering \citep{2017arXiv170100160G}. Generative models are also applicable to reinforcement learning, particularly model-based learning algorithms. A time series generative model, one which predicts future states of an environment given the current state of the environment and an agents actions, are of particular interest \citep{2017arXiv170100160G}. An agent may learn the best action to take in a given scenario by querying this generative model and choosing the one which outputs the best state of the world. Furthermore, generative models, especially GANs, have recently excelled at semi-supervised learning.

Proposed in the original GAN paper \citep{2014arXiv1406.2661G}, semi-supervised techniques used in conjunction with GANs have recently proven themselves as a reputable technique. The premise involves training the discriminator to classify images into $n + 1$ classes where $n$ is the number of class labels and the additional $1$ is the class of a fake image. This algorithm is able to obtain great performance on relatively small labeled datasets when compared to labeled datasets typically state of the art supervised algorithms \citep{2017arXiv170100160G}. Currently, the best performing semi-supervised GAN are feature matching GANs, where the new objective of the generative model is to minimize the distance between a layer $l$ of the discriminator \citep{2016arXiv160603498S}.

Many applications require multi-modal outputs. For example, predicting the next action of a self-driving car. Within this context, it is important the model not be trained by traditional methods, such as using mean squared error to minimize the distance between the expected and predicted actions. These models cannot be used in situations where there is more than one appropriate prediction. Generative modeling, especially GAN, allow machine learning to function the scenarios where one input corresponds to multiple acceptable outputs \citep{2017arXiv170100160G}.

Tasks which involve image generation, modification or translation are all highly applicable to generative models. Although GANs may be used to generate any type of data, images remain the most commonly used source of training data. Research conducted within the past three years demonstrates the capabilities of GANs to perform these tasks \citep{2017arXiv170100160G, 2016arXiv161102200T, 2014arXiv1411.1784M}. For example, GANs have recently been shown to be capable of translating aerial photographs into maps and sketches into realistic images \citep{2017arXiv170100160G}.

When compared to other generative models, especially deep generative models, GANs possess several advantages... quicker? no Markov chains?

\section{Background}
Generative adversarial networks may use any differentiable functions for the models; however, the optimal results are typically achieved when neural networks are used \citep{2014arXiv1406.2661G}. We first define a vector $z$ such that $z~p_z(z)$. This vector becomes the argument of the generator $G(z;\theta_g)$ where $G$ is a differentiable function and $\theta_g$ is its parameters. Another differentiable function $D(x;\theta_d)$ for the discriminator. This function takes as input a vector $x$, either from the output of $G$ or a sample from the training set, and is parametrized by $\theta_d$. The output of the discriminator is a scalar number representation the probability that the arguments were derived from the training data. In this situation, there exists a minimax game where $D$ attempts to maximize the probability of assigning a correct label and $G$ attempts to minimize $\log(1-D(G(z)))$. The networks $G$ and $D$ play a minimax game with the value function $V(D,G)$:

@@

\begin{figure}
\centering
\includegraphics[scale=0.45]{cumulative_gans}
\caption{The cumulative amount of papers written about generative adversarial networks. The graph shows the exponential growth of GANs within the past three years. Image maintained by Avinash Hindipur \citep{November2017GANZoo}.}
\label{fig:CumulativeNumberofGANPapersbyMonth}
\end{figure}

@@

\section{Latent Space}
The generator model of a GAN must internalize a representation of the data they are trained on. This quality is shared with variatinal autoencoders word linguistic models such as \textit{word2vec} \citep{2017arXiv171007035C}. One interesting characteristic about the latent space is that it is typically of lower dimension than the training data \citep{2017arXiv171007035C}. This forces the model to effectively represent the training data by extracting only useful features. As long as the effects of mode collapse are minimal, the latent space can be highly structured \citep{2017arXiv171007035C}. Often, interpolations allow for semantic transformations. For example, movement of the latent vector in certain directions may alter rotation, produce colour changes or see the addition of attributes such as eyeglasses \citep{2017arXiv171007035C}.

As mentioned in section \ref{sec:inferencegan}, it is possible to train an encoder to map a sample $\br{x}$ to a latent vector $\br{z}$. This is an extremely useful feature for exploring the latent space. One may discover \textit{concept vectors} such as "wearing glasses" or "wearing a hat" and apply them to the vectors associated with humans. This is a fascinating and practical property \citep{2017arXiv171007035C}.

@@

\section{Applications}
Although GANs are a relatively new topic, many different applications have been discovered since their inception. Several of the reoccurring themes are elaborated within this section.

@@

\subsection{Image Generation} \label{sec:img2img}
Image generation is extremely useful as many tasks intrinsically require the ability to generate realistic samples \citep{2017arXiv170100160G}. Most research conducted with respect the GANs attempts to improve the quality of the synthesized images \citep{2017arXiv171007035C}. As such, several useful and fascinating applications have already been found.

>>>>

\begin{abstract}
    Within the past decade, deep generative networks have often fallen short, especially when compared to deep discriminative models which are now being applied to a multitude of real world applications. However, since their inception in 2014, generative adversarial networks (GANs) have proved themselves as promising a research avenues for deep generative models. These networks now offer encouraging results for both unsupervised learning and generative models. The goal of this research paper is to present an overview of generative adversarial networks. Specifically, we will examine the motivation and challenges associated with GANs, their training process, architecture variations and some applications that are currently being explored.
\end{abstract}

\section{Introduction}
Deep learning has made astounding progress within the past decade. Discriminative models have already surpassed the abilities of humans to recognize patterns within certain domains \citep{2014arXiv1404.7828S}. These successes can attributed to vast, high dimension datasets in conjunction with large neural networks using linear activation functions, dropout regularization techniques and backpropagation to update the parameters \citep{2014arXiv1406.2661G}. However, deep learning possesses much more ambitious goals within the realm of unsupervised learning. As humans, we are able to understand the world around us with tremendous precision. It is easy to underestimate the complexity of the data we process to accomplish this feat. Although progress within the machine learning field is rapidly advancing, computers still have limited understanding of the data the process. Generative models, especially deep generative models such as GANs, offer promising results towards this formidable challenge \citep{genmodelingopenai}.

\subsection{Generative Modeling}
Generative models learn the joint probability distribution $p(\br{x},y)$ of an input $\br{x}$ and label $y$ whereas discriminative models directly learn the conditional probability $p(y|x)$. An example of of this process can be seen in figure \ref{pdf}. Generative models may be used as classifiers using Bayes rules to calculate the conditional probability $p(\br{x},y)$ which can then be used to make predictions \citep{NIPS2001_2020}. Although, generative models offer several other applications due to their knowledge of the underlying data distribution. Depending on the model, knowledge of the probability distribution can be created both explicitly and implicitly \citep{Goodfellow-et-al-2016}. Those which do not directly model a probability distribution offer mechanisms which require implicit knowledge of the underlying distribution, such as geminating a sample from that distribution \citep{Goodfellow-et-al-2016}.

\begin{figure}
\centering
\includegraphics[scale=0.7]{pdf}
\caption{The process density estimation of one-dimensional data and a Gaussian distribution \citep{2017arXiv170100160G}. Generative models take a dataset $D$, sourced from a distribution $p_{data}$ and create an estimate of the distribution $p_{model}$.}
\label{fig:pdf}
\end{figure}

The process of training a generative model is very similar to a discriminative model. Using large amounts of data collected from a specific domain, we train the generative model to generate data from that domain. As these models typically have fewer parameters than that the number of data samples, they are forced to internalize some representation of the data \citep{genmodelingopenai}. This process within the context of GANs is elaborated in section \ref{sed:latentspace}. Unlike supervised training, there exists no single desired output. The problem becomes defining a cost function which forces the generative model to produce data more like that of the target domain \citep{genmodelingopenai}.

\subsection{Generative Adversarial Networks}
Generative adversarial networks (GANs), first introduced in 2014 by Goodfellow \textit{et al.}, offer a new framework for estimating generative models with use of an adversarial process \citep{2014arXiv1406.2661G}. These models offer a clever approach to solving the aforementioned problem of explicitly defining a cost function. Rather than training a single model, a discriminator is introduced. These two networks are pitted against each other in a minimax, zero-sum game. The generative model $G$ attempts to produce data that resembles that of the training set while the discriminative model $D$ attempts to discriminate between a real and generated (fake) sample. To train this network, $G$ attempts to maximize the error rate of $D$ whereas $D$ attempts to minimize it. The goal of the GAN algorithm is to reduce the distance between the data and generated probability distributions. Backpropagation is used to update the parameters and to train each model. In the ideal case, $D$ is unable to distinguish the generated samples from the real samples and produces an error rate of $\frac{1}{2}$ \citep{2014arXiv1406.2661G}.

The adversarial process is often analogized to a counterfeiter trying to fool a detective \citep{2014arXiv1406.2661G}. The detective will receive both real and counterfeited money and will try and distinguish between the two. This leads to a competitive scenario where both parties gradually improve their abilities until the counterfeits are indistinguishable from the true currency.

\subsection{Motivations}
There exists several compelling reasons for studying generative modeling, especially GANs \citep{2017arXiv170100160G}. Generative models are extremely useful when your goal is understand the underlying data generating distribution. The ability to learn the joint probability distribution of high dimensional data is relevant to both applied math and engineering \citep{2017arXiv170100160G}. Furthermore, GAN promise many real world applications. Several of these opportunities are expanded upon in section \ref{sec:applications}.

\subsection{Related Work}
Until the advent of GANs, most deep generative models provided a parametric specification of an probability distribution function \citep{2014arXiv1406.2661G}. These models could be trained using maximum likelihood estimation which often required gradient estimation. The most successful of these models is the deep Boltzmann machine \citep{2014arXiv1406.2661G}. The difficulties associated with gradient estimation motivated the development of \texit{generative machines}, models which implicitly represent the joint probability distribution by generating samples from that distribution rather than explicitly rendering a probability distribution.

Deep generative stochastic networks are an example of a generative machine. This model, developed by Bengio \textit{et al.} in 2013 \citep{2013arXiv1306.1091B}, was able to use backpropagation during training; however, still made use Markov chains which introduced instability problems. GANs extend the ideas presented within generative stochastic networks by eliminating the use of Markov chains \citep{2014arXiv1406.2661G}.

\section{Background}
GANs may use any differentiable functions for the models; however, the optimal results are typically achieved when neural networks are used \citep{2014arXiv1406.2661G}. We first define a vector $z$ such that $z~p_z(z)$. This vector becomes the argument of the generator $G(z;\theta_g)$ where $G$ is a differentiable function and $\theta_g$ is its parameters. Another differentiable function $D(x;\theta_d)$ for the discriminator. This function takes as input a vector $x$, either from the output of $G$ or a sample from the training set, and is parametrized by $\theta_d$. The output of the discriminator is a scalar number representation the probability that the arguments were derived from the training data. In this situation, there exists a minimax game where $D$ attempts to maximize the probability of assigning a correct label and $G$ attempts to minimize $\log(1-D(G(z)))$. The networks $G$ and $D$ play a minimax game with the value function $V(D,G)$:

@@

\begin{figure}
\centering
\includegraphics[scale=0.43]{cumulative_gans}
\caption{The cumulative amount of papers written about generative adversarial networks. The graph shows the exponential growth of GANs within the past three years. Image maintained by Avinash Hindipur \citep{November2017GANZoo}.}
\label{fig:CumulativeNumberofGANPapersbyMonth}
\end{figure}

@@

\section{Latent Space} \label{sed:latentspace}
The generator model of a GAN must internalize a representation of the data they are trained on. This quality is shared with variatinal autoencoders word linguistic models such as \textit{word2vec} \citep{2017arXiv171007035C}. One interesting characteristic about the latent space is that it is typically of lower dimension than the training data \citep{2017arXiv171007035C}. This forces the model to effectively represent the training data by extracting only useful features. As long as the effects of mode collapse are minimal, the latent space can be highly structured \citep{2017arXiv171007035C}. Often, interpolations allow for semantic transformations. For example, movement of the latent vector in certain directions may alter rotation, produce colour changes or see the addition of attributes such as eyeglasses \citep{2017arXiv171007035C}.

As mentioned in section \ref{sec:inferencegan}, it is possible to train an encoder to map a sample $\br{x}$ to a latent vector $\br{z}$. This is an extremely useful feature for exploring the latent space. One may discover \textit{concept vectors} such as "wearing glasses" or "wearing a hat" and apply them to the vectors associated with humans. An example of this process can be seen in figure \ref{fig:conceptvector}. This is a fascinating and practical property \citep{2017arXiv171007035C}.

@@

\section{Applications} \label{sec:applications}
Although GANs are a relatively new topic, many different applications have been discovered since their inception. Several of the reoccurring themes are elaborated within this section.

\subsection{Image Generation} \label{sec:img2img}
Image generation is extremely useful as many tasks intrinsically require the ability to generate realistic samples \citep{2017arXiv170100160G}. Moreover, GANs posses the ability to model multi-modal outputs (see figure \ref{fig:multimodal}) \citep{2017arXiv170100160G}. Most research conducted with respect the GANs attempts to improve the quality of the synthesized images \citep{2017arXiv171007035C}. As such, several useful and fascinating applications have already been found.

\begin{figure}
\centering
\includegraphics[scale=0.7]{multimodal}
\caption{An example of scenario where the ability to model multi-modal data is important. This image demonstrates the process of predicting the next video frame given the previous one. A generator trained with mean squared error (middle) produces an image which averages each possible future. Using an adversarial process (right), the image produced corresponds with the most likely scenario. Example retrieved form Lotter \texit{et al.} \citep{2015arXiv151106380L}.}
\label{fig:multimodal}
\end{figure}

@@
